---
title: "Loan Analysis"
output: 
    flexdashboard::flex_dashboard:
        orientation: columns
        vertical_layout: fill
        self_contained: true
        social : ["twitter","facebook","menu"]
        source_code : embed
---

```{r load-packages}
library(neuralnet)
library(rbokeh)
library(psych)

```


```{r load-dataset}

loandata <- read.csv("C:\\Users\\teeja\\Desktop\\CleanGithubProjects\\scm651\\hw4_gb3\\scm651_homework_4_universal_bank.csv", na.strings =c(""))
#head(loandata)

```



# Data Decription
## {.tabset}
### Columns Descrption

#### Background
Using the Universal Bank data, determine the factors which influence whether a customer
takes out a loan

#### Universal Bank Data Fields
1. ID : unique identifier
1. Personal Loan :  did the customer accept the personal load offered (1=Yes, 0=No)
1. Age : customer’s age
1. Experience :  number of years of profession experience
1. Income : annual income of the customer ($000)
1. Zip code:  home address zip code
1. Family : family size of customer
1. CCAvg : average spending on credit cards per month ($000)
1. Education:  education level (1) undergraduate, (2) graduate, (3) advanced/professional
1. Mortgage :  value of house mortgage ($000)
1. Securities :  does the customer have a securities account with the bank? (1=Yes, 0=No)
1. CDAccount : does the customer have a certificate of deposit with the bank? (1=Yes, 0=No)
1. Online : does the customer use Internet banking facilities (1=Yes, 0=No)
1. CreditCard : does the customer use a credit card issued by Universal Bank? (1=Yes, 0=No)

### Summary
```{r summary-data}
summary(loandata)
```

### Data Structure
```{r}
str(loandata)
```




```{r}
### Data Transformation
#loandata$CustomerID < as.factor(loandata$CustomerID)
originaldata <- loandata

# # tranform loan data
# loandata$PersonalLoan <- as.factor(loandata$PersonalLoan)
# loandata$Income <- as.numeric(loandata$Income)
# loandata$ZIP.Code <- as.factor(loandata$ZIP.Code)
# loandata$Education <- as.ordered(loandata$Education)
# loandata$Mortgage <- as.numeric(loandata$Mortgage)
# loandata$SecuritiesAccount <- as.factor(loandata$SecuritiesAccount)
# loandata$CDAccount <- as.factor(loandata$CDAccount)
# loandata$Online <- as.factor(loandata$Online)
# loandata$CreditCard <- as.factor(loandata$CreditCard)
# 
# str(loandata)

```

### Scatter Plot

```{r}
p = loandata  %>% figure(title = "Scatter Plot", width = 1000) %>% ly_points(x=Experience, y=Income, size = CCAvg, color = PersonalLoan,
                                                               hover = list(ZIP.Code, CDAccount,Online, PersonalLoan) )
p

```


### Scater Matrix
```{r}
pairs.panels(loandata[,-c(1,6)])

```



### Age and Experience

```{r}
p =loandata[,c(3,4)] %>% figure(title = "Age and Experience") %>%
      ly_points(Age,Experience, hover = list(Age, Experience)) 

p

```





# Logit

## {.tabset}

### Model 1


```{r split-data}

ind <- sample(2, nrow(loandata), replace = TRUE, prob = c(0.7, 0.3))
train <- loandata[ind==1,]
test <- loandata[ind==2,]

# remove customer ID

train <- train[,-1]
test <- test[,-1]

# str(train)
#str(test)
#dim(train)
#dim(test)

```




```{r}

#remove zip code

model1 <- glm(PersonalLoan ~ .,family = binomial("logit"), data= train[,-5])

summary(model1)



```

### Model 2
```{r}

#remove zip code

model2 <- glm(PersonalLoan ~ .,family = binomial("logit"), data= train[, -c(2,5)])

summary(model2)

```


### Model 3

```{r}
#remove zip code, Age, Experience

model3 <- glm(PersonalLoan ~ .,family = binomial("logit"), data= train[,-c(2,3,5,9)])

summary(model3)

```


### ANOVA

```{r}
anova(model3, test="Chisq")
```


### Model Accuracy

```{r}

prediction <- predict.glm(model3, newdata = test[,-c(1,2,3,5,9)])

prediction <- ifelse(prediction > 0.5, 1, 0)

error <- mean(prediction != test$PersonalLoan)


prediction2 <- predict.glm(model2, newdata = test[,-c(1,2,5)])

prediction2 <- ifelse(prediction2 > 0.5, 1, 0)

error2 <- mean(prediction2 != test$PersonalLoan)


prediction1 <- predict.glm(model1, newdata = test[,-c(1,5)])

prediction1 <- ifelse(prediction1 > 0.5, 1, 0)

error1 <- mean(prediction1 != test$PersonalLoan)


cat("Model 1 Accuracy : ", 1-error1, '\n\n')
cat("Model 2 Accuracy : ", 1-error2, '\n\n')
cat("Model 3 Accuracy : ", 1-error, '\n')
 
```




# Probit

## {.tabset}

### Model 1




```{r}

#remove zip code

model1 <- glm(PersonalLoan ~ .,family = binomial("probit"), data= train[,-5])

summary(model1)



```

### Model 2
```{r}

#remove zip code

model2 <- glm(PersonalLoan ~ .,family = binomial("probit"), data= train[, -c(2,5)])

summary(model2)

```


### Model 3

```{r}
#remove zip code, Age, Experience

model3 <- glm(PersonalLoan ~ .,family = binomial("probit"), data= train[,-c(2,3,5,9)])

summary(model3)

```


### ANOVA

```{r}
anova(model3, test="Chisq")
```


### Model Accuracy

```{r}

prediction <- predict.glm(model3, newdata = test[,-c(1,2,3,5,9)])

prediction <- ifelse(prediction > 0.5, 1, 0)

error <- mean(prediction != test$PersonalLoan)


prediction2 <- predict.glm(model2, newdata = test[,-c(1,2,5)])

prediction2 <- ifelse(prediction2 > 0.5, 1, 0)

error2 <- mean(prediction2 != test$PersonalLoan)


prediction1 <- predict.glm(model1, newdata = test[,-c(1,5)])

prediction1 <- ifelse(prediction1 > 0.5, 1, 0)

error1 <- mean(prediction1 != test$PersonalLoan)


cat("Model 1 Accuracy : ", 1-error1, '\n\n')
cat("Model 2 Accuracy : ", 1-error2, '\n\n')
cat("Model 3 Accuracy : ", 1-error, '\n')
 
```



# Moderating Probit

##{.tabset}

### Probit Model 1

```{r}

l1_model <- glm(PersonalLoan ~ (Income + Family + CCAvg + Education + Mortgage + SecuritiesAccount + CDAccount + Online + CreditCard)^2 ,
             data = train, family = binomial("probit"))

summary(l1_model)

```


### Probit Model 2

```{r}
 
l2_model <- glm(PersonalLoan ~ Income + Family + CCAvg + Education + Mortgage + SecuritiesAccount  + Online + CreditCard + (Income * Family) +
               (Income * CCAvg) + (Income + Education) + (Family + Education) + (Family * SecuritiesAccount) + (CCAvg * Education) +
               (Online * CreditCard),
             data = train, family = binomial("probit"))
summary(l2_model)
```


### Probit Model 3

```{r}

 
l3_model <- glm(PersonalLoan ~ Income + Family + CCAvg + Education  + (Income * Family) +
               (Income * CCAvg)  + (CCAvg * Education) ,
             data = train, family = binomial("probit"))

summary(l3_model)

```



### Moderating Accuracy

```{r}
#model 1

prediction <- predict.glm(l1_model, newdata = test[,-c(1,5)])

prediction <- ifelse(prediction > 0.5, 1, 0)

error <- mean(prediction != test$PersonalLoan)


# model 2

prediction2 <- predict.glm(l2_model, newdata = test[,c("Income", "Family", "CCAvg", "Education", "Mortgage", "SecuritiesAccount","Online", "CreditCard" )])

prediction2 <- ifelse(prediction2 > 0.5, 1, 0)

error2 <- mean(prediction2 != test$PersonalLoan)


prediction3 <- predict.glm(l3_model, newdata = test[,c("Income","Family", "CCAvg", "Education" )])

prediction3 <- ifelse(prediction3 > 0.5, 1, 0)

error3 <- mean(prediction1 != test$PersonalLoan)


cat("Model 1 Accuracy : ", 1-error, '\n\n')
cat("Model 2 Accuracy : ", 1-error2, '\n\n')
cat("Model 3 Accuracy : ", 1-error3, '\n')

```








# Moderating Logit

##{.tabset}

### Logit Model 1

```{r}

l1_model <- glm(PersonalLoan ~ (Income + Family + CCAvg + Education + Mortgage + SecuritiesAccount + CDAccount + Online + CreditCard)^2 ,
             data = train, family = binomial("logit"))

summary(l1_model)

```


### Logit Model 2

```{r}
 
l2_model <- glm(PersonalLoan ~ Income + Family + CCAvg + Education + Mortgage + SecuritiesAccount  + Online + CreditCard + (Income * Family) +
               (Income * CCAvg) + (Income + Education) + (Family + Education) + (Family * SecuritiesAccount) + (CCAvg * Education) +
               (Online * CreditCard),
             data = train, family = binomial("logit"))
summary(l2_model)
```


### Logit Model 3

```{r}

 
l3_model <- glm(PersonalLoan ~ Income + Family + CCAvg + Education  + (Income * Family) +
               (Income * CCAvg)  + (CCAvg * Education) ,
             data = train, family = binomial("logit"))

summary(l3_model)

```



### Moderating Accuracy

```{r}
#model 1

prediction <- predict.glm(l1_model, newdata = test[,-c(1,5)])

prediction <- ifelse(prediction > 0.5, 1, 0)

error <- mean(prediction != test$PersonalLoan)


# model 2

prediction2 <- predict.glm(l2_model, newdata = test[,c("Income", "Family", "CCAvg", "Education", "Mortgage", "SecuritiesAccount","Online", "CreditCard" )])

prediction2 <- ifelse(prediction2 > 0.5, 1, 0)

error2 <- mean(prediction2 != test$PersonalLoan)


prediction3 <- predict.glm(l3_model, newdata = test[,c("Income","Family", "CCAvg", "Education" )])

prediction3 <- ifelse(prediction3 > 0.5, 1, 0)

error3 <- mean(prediction3 != test$PersonalLoan)


cat("Model 1 Accuracy : ", 1-error, '\n\n')
cat("Model 2 Accuracy : ", 1-error2, '\n\n')
cat("Model 3 Accuracy : ", 1-error3, '\n')

```


# Final Model
##
### Model

```{r}

model  <- glm(PersonalLoan ~ Income + Family + CCAvg + Education  + (Income * Family) +
               (Income * CCAvg)  + (CCAvg * Education) ,
             data = loandata, family = binomial("logit"))

summary(model)



```



# Neural Network

## {.tabset}

### Mode 



```{r}

model1_nn <- neuralnet(PersonalLoan ~ Income + Family + CCAvg + Education  , hidden = 3,
                        lifesign = "minimal",linear.output = FALSE,threshold = 0.1 , data = train )



```



### Neural Networks


```{r}

model1_nn$result.matrix

plot(model1_nn)


```



### Accuracy

```{r}

test_result <- compute(model1_nn,test[,c("Income","Family", "CCAvg", "Education" )] )

prediction <- test_result$net.result


prediction3 <- ifelse(prediction > 0.5, 1, 0)

error <- mean(prediction3 != test$PersonalLoan)

cat("Accuracy : ", 1-error)


```



### Result Matrix

```{r}
model1_nn$result.matrix

```


# Analysis in Excell

#### [Analysis in Excel](https://sumailsyr-my.sharepoint.com/:x:/g/personal/toabdula_syr_edu/Ed793jjWAg1NrkXHHfbzA90BewwooVhx_3kTZzHe3BU7lA?e=MqZmqz)


