---
title: "Introduction to R Programming"
author: "Tajudeen Abdulazeez"
date: "5/19/2023"
output: html_document
link-citations: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction to R

RStudio requires R 2.11.1 (or higher).
Download and install R first
Then download and install R studio

We will be going through this step by step in the session.

### GETTING R AND R STUDIO
Links:

- https://www.r-project.org/

- https://www.rstudio.com/

- https://www.rstudio.com/resources/webinars/getting-your-data-into-r/

- https://www.rstudio.com/products/rstudio/download/


### USING R

USING KNOWN AND FREE DATASETS

R offers many datasets that can be used to practice with.
R dataset information and examples are located here:
https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/00Index.html
R Studio installs with the R “datasets” Package. 
To use it, type (or include in your code) the statement:
library(datasets)

```{r}
library(datasets)

head(cars)
```

```{r}
head(iris)
```


#### PULLING DATA INTO YOUR R PROGRAM

```{r}
(x <- 1:10)  #This will set x as an array from 1 to 10 and will print it. 
(y <- 21:30)

plot(x,y, xlab="x label", ylab="HEIGHT")

print("Hello")
(x[3])
print(x[0])

(p <- rep(3,9))
(length(p))

(mean(p))

```




#### REP AND SEQ

```{r}
(q<-rep("TRUE",3))
(y<-seq(from=1, to=9, by=2))
(z<-seq(11,23,3))
(x<-c(2.2, 3.0, 6.7))
w<-c(3.1, 5.5, 8.7, 9.9)
w[3]
min(w)
max(w)
sort(w)
(s1<-rep(w,times=2))
(s2 <- rep(w,each=2))

```


#### CREATE A MATRIX: RBIND AND CBIND

```{r}
d1 <- c(0, 1, 3, 5, 9)
d2<- c(3, 6, 0, 34, 67)
(rbind(d1,d2))
(cbind(d1,d2))

```


#### DATAFRAMES

```{r}
#dataframes
#a list with elements of diff modes (types)
(df <- list("apple", 7, TRUE, "fred", 5.5))

(mode(df))

#sapply - apply a particpular function to #each element
(sapply(df,mode))

```


#### USING “C” (CONCAT) VERSUS “LIST”
Creating an array using “c” will require that all items in the array are the SAME TYPE. 
Lists do not have to be of the same type

```{r}
#c vs list
(f <- c(5, 7, "dog", 9))

# Because concat defaults it will make all #chars
#list allows a collections of different modes
#or types
(g <- list(5, 7, "dog", 9))
(sapply(f,mode))
(sapply(g,mode))

```


#### EXAMPLE: MAKING A DATAFRAME

```{r}
(df1 <- data.frame(book=c("B1", "B2", "B3"), author=c("A1", "A2", "A3"), year=c(1950, 2005, 1999), used=c(TRUE, FALSE, TRUE)))


(df1[1,2])
#dataframes can be made column by column

```

#### DATAFRAME LARGER EXAMPLE: AARDVARKS

```{r}
aardvarkdata<-data.frame(individual=1:10, country=c(rep("Kenya",6), rep("Tanzania",4)), litter=c(2, 3, 4, 2, 3, 4, 2, 3, 1, 3))

print(aardvarkdata)

aardvarkdata$litter
aardvarkdata$country
aardvarkdata[,3]

# All rows, columns 1 and 2
aardvarkdata[,1:2]

```


#### ADD A NEW COLUMN (MEANWEIGHT) TO DATAFRAME

```{r}
aardvarkdata$meanWeight<-c(30, 29, 31, 32, 34, 32, 34, 45, 34, 23)
print(aardvarkdata,row.names = FALSE)

```


#### PHYSICALLY IMPORTING DATA IN R STUDIO

```{r}
MyData <- read.csv(file="/Users/teejay/dev/courses/machine-learning-in-r/data/SalesData.csv")
(head(MyData)) #prints the first 6 rows and the headers
#Can also use read.table(read.csv(file="C:/Users/Ami/Desktop/nySMALL.csv", sep="\t", header=TRUE)
#Can Also allow the User to Choose the file as the R code runs:
#AirData <- read.table(file.choose(), sep="\t", header=TRUE)
#It is also possible to directly read the HTML code from a website
#sitedata <- read.scv("http://www.mathandstatistics.com")
Store <- MyData$Store

Mon <- MyData$Mon


library(lattice)
histogram(Mon)
tail(MyData)
nrow(MyData)
ncol(MyData)

```




#### GETTING NEW PACKAGES

- install.packages('package_name')


#### USING R DATASETS: AIRQUALITY

```{r}
#Air Quality
mydata<-airquality

(head(mydata))

(tail(mydata))

```


#### WRITING TO A DATAFILE FROM R

```{r}
MyData <- read.csv(file="/Users/teejay/dev/courses/machine-learning-in-r/data/SalesData.csv")
(head(MyData))

Region <- MyData$Region
write.csv(Region, file="/Users/teejay/dev/courses/machine-learning-in-r/data/NewRFile.txt")

```


#### LOOKING AT DATA

-  Looking at raw data: head, tail, class, levels, and view

-  What type of data object is the data stored in: vector/matrix, dataframe, list, etc.

-  What are properties of the data: nrow(), ncol(), length(), names(), str(), dplyr::glimpse()

-  Visualize (basic) the data: plot(), identity()

-  Summarize the data: xtab()


#### LOOKING AT THE DATA

```{r}

#Must first install.packages("dplyr") on console
library(dplyr)

AirData <- airquality
(head(AirData))
(tail(AirData))
(class(AirData))
(length(AirData))  # number of columns
(ncol(AirData))  # same as length
(names(AirData))  #col names


(nrow(AirData))
((AirData[10,4]))  #[rows,columns]
#(View(AirData))  #"V" must be uppercase

#if larger than 1000 rows, view will only #write first 1000
(str(AirData))   #summarize the data
(levels(AirData))  #
#(glimpse(AirData))


```


#### STARTING TO CLEAN UP THE DATA
```{r}
#Change Month to variable to a factor (not an int)
AirData$Month <- factor(AirData$Month)

(levels(AirData$Month))
#Change the names of the levels
levels(AirData$Month)<- list(May="5", June="6", July="7", Aug="8", Sept="9")
(levels(AirData$Month))
#(glimpse(AirData))

```


#### VISUALIZING

```{r}
plot(AirData$Ozone,AirData$Temp)

```


#### USING IDENTIFY
```{r}
plot(AirData$Ozone,AirData$Temp)
## identify makes the plot interactive - as you #click
# points are saved. Must press "finish" at the #top of the plot
# identify will then label the dta points clicked.
idx<-identify(AirData$Ozone, AirData$Temp)

```


#### IDENTIFY WITH IMPROVED LABELS
```{r}
## Get the Day for each datapoint
# Here, first paste together the char #version of the day with _ and the char
#version of the month. Use this to label #all points clicked. 
plot(AirData$Ozone,AirData$Temp)
idx<- identify(AirData$Temp, labels = paste(as.character(AirData$Day),"-",as.character(AirData$Month)), plot=TRUE)


```


#### XTAB
```{r}
## xtabs is a function that will investigate
#Is there an even distribution of data labels
## It is a formula/function so we use the ~
## It will COUNT the number of investigates for each label in dataset
xtabs(~Month, data=AirData)

```



#### References 

 R Refresher  by Dr. Ami Gates
Associate Professor, Georgetown University
drgates@georgetown.edu



